Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas

### 2021-12-19 analysis/paper/paper.Rmd

```{r getdata}
targets::tar_load(analysis_data, store = here::here("_targets"))
targets::tar_load(grid, store = here::here("_targets"))
targets::tar_load(deploy_table, store = here::here("_targets"))
```
Data used in our study came from bio-loggers deployed on bearded, ribbon, and
spotted seals in the Bering, Chukchi, and Beaufort seas by multiple
organizations as part of collaborative investigations during 2005 and 2020.
Seals were captured using nets and instrumented during ship-based capture trips
and from coastal Alaska communities. Ship-based capture events occurred during
spring near the southern ice edge in the Bering Sea from 20xx to 20xx (maybe
provide sample sizes for each species). Land-based capture events occurred from
May to October, generally between the coastal communities of Scammon Bay,
Alaska, in the Bering Sea, and Utqiaġvik, Alaska, near the junction of the
Chukchi and Beaufort seas, from 2005 to 2020. We refer readers to the primary
literature for detailed capture and bio-logger attachment methods (Supplemental
Material, Table).

@@

Tags that fall off due to molt or mortality and remain on ice or land can still
send data to satellites; i.e., the bio-logger will be on ice or land and dry, therefore, it will record and transmit data suggesting the seal is hauled out. As such, end times of each deployment were identified by examining
bio-logger locations and dive behavior to determine when bio-loggers ceased providing data
consistent with seal behavior. Data outside of the deployment start and end
times were discarded prior to analysis. After approximately 9 months, 7 devices
deployed on the rear flipper of bearded seals reported implausible hourly percent-dry data.
All data after the first instance of unrealistic values were censored from this
analysis. Figure \@ref(fig:dataCal) shows a calendar-like distribution of all
haul-out data across the study season for each species. Observations for ribbon
and spotted seals are concentrated in the months of May and June due to the
timing of deployment (April and May) and the timing of molt (May and June).
During molt, seals (and their attached bio-loggers) spend more time out of the
water and more data are transmitted. Molt timing also impacts when many
deployments end as any bio-loggers adhered to the hair will fall off. Bearded seal
deployments are more evenly spread with fewer observations in May and June. The
majority of bearded seal deployments start later in the summer and by May bio-loggers
have either fallen off or the batteries are depleted.

@@

ggplot(data = seal_hours_tbl) + 
  geom_point(aes(x = col_group, y = row_group, color = n, size = n), stroke = 0, shape = 15) +
  scale_size(range = c(1,3.5), breaks = c(100,500,1000,1500)) +
  scale_color_gradient(low = "#6baed6", high = "#08306b") +
  scale_x_continuous(breaks = c(1,4,8,12,16),
                     labels = c("March", "April", "May", "June", "July")) +
  scale_y_discrete(limits = rev) +
  guides(color = "none",
        size = guide_legend(override.aes = list(color = "#2171b5"),
                            title.position = "bottom", title.hjust = 0.5)
        ) +
  facet_grid(species ~ .) + 
  theme(
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(x = "", y = "", color = "seal-hours",
       size = "seal hours of bio-logger data") +
  ggtitle("Density of haul-out behavior observations varies from March through July",
          subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")
```

Haul-out behavior data was recorded and transmitted via the Argos satellite
network as hourly percent-dry timelines. For each hour of a day, the wet/dry
sensor was polled by the tag firmware every few seconds and a percent of the
hour in a dry state was calculated (Figure \@ref(fig:examplePlot)). Hourly
percent-dry data were usually rounded to the nearest 10 percent inclusive of 0
and 100 percent and additional values at 3 and 98 percent. This compression
resulted in additional data transmission as each message consisted of two
complete 24-hour records. Bio-loggers were, generally, deployed and programmed
in a manner to maximize data transmission during the spring pupping and molting
period, however, hourly percent-dry data for days during a deployment were not
successfully transmitted. This is due to a variety of factors including
satellite coverage, tag availability (i.e. tags mounted to the rear flipper are
not available to the surface while the seal is at sea), tag performance, duty
cycling, and extra-terrestrial atmospheric interference.

@@

p <- p + facet_grid(year~month, scales = "free_x")
p <- p + scale_x_continuous(breaks = c(5,15,25), expand = c(0,0)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  theme(panel.spacing = unit(0,"line")) +
  ggtitle("Observation records are incomplete because not all data are transmitted",
          subtitle = "percent-dry data for a single ribbon seal across two years") +
  xlab("day of month") + ylab("hour (UTC)") 
p <- p + 
  geom_point(aes(x = day, y = hour), data = missing_hours, 
             color = "black", shape = 4, size=1)
p
```

Of key interest in this study is the relationship between haul-out behavior and
environmental covariates that vary with space and time. We explored the use of a
continuous-time correlated random walk [@devins.johnson2008] movement model to
predict locations at specific times. However, the sparse nature of data from
some bio-loggers, especially those mounted to the rear flipper, resulted in poor
modeling performance or convergence issues. For this study, we calculated a
weighted average daily location where the inverse of the estimated Argos or
FastLoc GPS location error was used for the weight. Each Argos location estimate
was assigned an error radius based on either the categorical location quality (
_3_=250m, _2_=500m, _1_=1500m, _0_=2500m, _A_=2500m, _B_=2500m) or, when
available, the estimated error radius from the Argos Kalman filter algorithm.
Location estimates from FastLoc GPS were all assigned an error radius of 50m.
Any days where haul-out observations were present but location data were missing
we used the last calculated weighted average daily location, and any days where
the location intersected with land were removed from the data set. We recognized
that bearded and spotted seals haul out on land. However, assessing the
relationship between haul-out behavior and environmental covariates and their
availability for aerial surveys on land was outside the scope of this study.
Additionally, any daily locations on land are likely more reflective of
coordinate averaging and measurement error instead of actual use of coastal
features. Figure \@ref(fig:dataMap) shows the spatial distribution weighted
locations with available haul-out behavior data for each species across the
study area.

@@

In addition to sex and age class, we analyzed variables that might help explain
variation in haul-out probabilities. These included day-of-year (for seasonal
effects) and local solar hour (for diurnal effects). Solar hour was calculated
using the {solaR} package [@perpinan2012] within the R statistical environment
[@rcoreteam2021] using the daily locations. We included several meteorological
variables shown to affect haul-out behavior in other Arctic pinnipeds
[@reder2003; @udevitz2009; @perry2017]. In particular, we linked the weighted
average daily locations from our bio-loggers to key weather values from the
North American Regional Reanalysis (NARR) model produced by the National Centers
for Environmental Prediction [@mesinger2006]. The NARR model assimilates
observational data to produce a long-term picture of weather over North America.
Numerous weather variables are made available across the region 8 times daily.
For this study, NARR weather values were subset to the extent of our study area
over the Bering and Chukchi seas at 3-hr intervals based on the native grid
resolution of 32 km (1024 sq. km). The following meteorological variables were
interpolated and assigned to seal locations using a bilinear method: 1) air
temperature at 2m above the earth's surface, 2) wind consisting of northerly and
easterly vector components converted to wind speed using the Euclidean norm, 3)
barometric pressure at sea level, and 4) precipitation (Table
\@ref(tab:tblCovariates)).

@@

covariate_tbl <- tibble::tribble(
  ~Covariate, ~Type, ~Source, ~Description,
  "Age-sex class",   "Categorical", "Field Assessment", 
  "young-of-the-year, subadult, adult male and adult female",
  "Hour", "Continuous; Fourier basis", "Bio-logger", 
  "local solar hour using 6 variables of a Fourier-series basis",
  "Day",   "Continuous", "Bio-logger", 
  "linear, quadratic, and cubic effects of day-of-year",
  "Precip", "Continuous", "NARR", 
  "convective precipitation (kg/m2)",
  "Pressure", "Continuous", "NARR", 
  "atmospheric pressure at sea level (kPa)",
  "Temp", "Continuous", "NARR",
  "air temperatures at 2m above the earth’s surface",
  "Wind", "Continuous", "NARR", 
  "northerly and easterly vector components for wind converted into a single wind speed via the Euclidean norm",
  "Northing", "Continuous", "Bio-logger","latitude divided by the mean latitude across all locations (for bearded seals only)"
)

covariate_tbl %>% 
  flextable() %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  set_table_properties(width = .95, layout = "autofit")
```

@@

Bearded seal haul-out probability appears most influenced by wind
($F_{`r paste(bearded_type3$wind$Num.df, bearded_type3$wind$Den.df, sep=",")`}$
= `r bearded_type3$wind$F.value`; $p$ = `r bearded_type3$wind$Prob.F`)
and temperature
($F_{`r paste(bearded_type3$temp2$Num.df, bearded_type3$temp2$Den.df, sep=",")`}$
= `r bearded_type3$temp2$F.value`; $p$ =
`r bearded_type3$temp2$Prob.F`). Additionally, wind chill was also a
significant influence
($F_{`r paste(bearded_type3$temp2_wind$Num.df, bearded_type3$temp2_wind$Den.df, sep=",")`}$
= `r bearded_type3$temp2_wind$F.value`; $p$ =
`r bearded_type3$temp2_wind$Prob.F`). Any influence of barometric
pressure or precipitation is less apparent.

@@

Compared to ribbon seals, spotted seals showed a longer spring haul-out season
that is less intensely centered on solar noon (Figure \@ref(fig:spottedHOCal)).
Adults of both sexes spend considerable time in April, May and June hauled out.
In contrast to ribibon seals, adult spotted seal males have a more protracted
haul-out season compared to females, and more time out of the water in June and
July (Figures \@ref(fig:ribbonHOCal) and \@ref(fig:spottedHOCal)). As with
ribbon seals, the young-of-the-year records begin after weaning and the model
predictions demonstrate the ontogeny of in-water activities (e.g. diving,
foraging) in May.

@@

We modeled data from bio-loggers deployed on bearded, ribbon, and spotted seals
to examine factors affecting haul-out behavior on sea ice in the Bering and
Chukchi seas. Our analysis shows all three species of seal haul out
progressively more through the spring and peak near mid-May to early June before
declining again. This pattern aligns well with what has been previously
documented and confirms our haul-out data are likely capturing population-level
behavioral patterns. Seals preferentially haul out on ice shortly after solar
noon, which allows seals to maximize absorption of solar radiation thought to
facilitate the molting process [@feltz1966]. Interestingly, bearded seals appear
to have two peaks in haul-out activity within a day, one shortly after solar
noon, and one centered near solar midnight. A similar bi-modal pattern has been
seen in ringed seals [@vonduyke2020] and suggests that bearded and ringed seals
may be operating under different constraints than ribbon and spotted seals.
Bearded and ringed seals are distributed across higher latitudes and the
extended daylight hours may allow more flexibility in optimizing resting periods
with foraging. Other factors such as polar bear predation (which is rare for
ribbon and spotted seals in the Bering Sea) may also explain differing haul-out
patterns. Bearded seals also showed a more protracted season of haul-out
behavior compared to ribbon and spotted seals. This aligns with findings from
Thometz et al. [@thometz2021] who observed a mean molting period of 119$\pm$2
days and a relatively stable resting metabolic rate during that time. While
ribbon seals were not considered in the same study, spotted and ringed seals
underwent molt periods of 33$\pm$4 and 28$\pm$6 days and exhibited increased
resting metabolic rates.

@@

We also investigated the influence of environmental covariates on haul-out
probabilities, including wind speed, temperature, barometric pressure,
precipitation, and wind chill. These have been investigated for walruses (e.g.
Udevitz et al. [-@udevitz2009]) and a few select studies of ice-associated seals
[@hamilton2018, @perry2017]. Ribbon seals seemed to be the most influenced by
weather, with wind, temperature, barometric pressure, and precipitation all
being important components of the model. Spotted seals were most affected by
wind and barometric pressures. And, for bearded seals, the model indicated wind
and temperature had the greatest impact. In general, and as might be expected,
seals were more likely to haul-out when daily temperatures were warmer, winds
speeds were lower, barometric pressure was higher, and precipitation was lower.
These weather conditions are likely most beneficial for energetic considerations
and may also provide the best situation for predator detection. These results
highlight the importance for haul-out behavior models (and subsequent
calculations of availability during surveys) to incorporate weather covariates.

>>>>

```{r getdata}
targets::tar_load(analysis_data, store = here::here("_targets"))
targets::tar_load(grid, store = here::here("_targets"))
targets::tar_load(deploy_table, store = here::here("_targets"))
```
Data used in our study came from bio-loggers deployed on bearded, ribbon, and
spotted seals in the Bering, Chukchi, and Beaufort seas by multiple
organizations as part of collaborative investigations from 2005 through 2020.
Seals were captured using nets and bio-loggers attached during ship-based
capture trips and from coastal Alaska communities. Ship-based capture events
occurred during spring near the southern ice edge in the Bering Sea. Land-based
capture events occurred from May to October, generally between the coastal
communities of Scammon Bay, Alaska, in the Bering Sea, and Utqiaġvik, Alaska,
near the junction of the Chukchi and Beaufort seas, from 2005 to 2020. We refer
readers to the primary literature for detailed capture and bio-logger attachment
methods (Supplemental Material, Table).

@@

Tags that fall off due to molt, attachment failure, or seal mortality and remain
on ice or land can still send data to satellites; i.e., the bio-logger will be
on ice or land and dry, therefore, it will record and transmit data suggesting
the seal is hauled out. As such, end times of each deployment were identified by
examining bio-logger locations and dive behavior to determine when bio-loggers
ceased providing data consistent with seal behavior. Data outside of the
deployment start and end times were discarded prior to analysis. After
approximately 9 months, 7 devices deployed on the rear flipper of bearded seals
reported implausible hourly percent-dry data. All data after the first instance
of unrealistic values were censored from this analysis. Figure
\@ref(fig:dataCal) shows a calendar-like distribution of all haul-out data
across the study season for each species. Observations for ribbon and spotted
seals are concentrated in the months of May and June due to the timing of
deployment (April and May) and the timing of molt (May and June). During molt,
seals (and their attached bio-loggers) spend more time out of the water and more
data are transmitted. Molt timing also impacts when many deployments end as any
bio-loggers adhered to the hair will fall off. Bearded seal deployments are more
evenly spread with fewer observations in May and June. The majority of bearded
seal deployments start later in the summer and by May bio-loggers have either
fallen off or the batteries are depleted.

@@

ggplot(data = seal_hours_tbl) + 
  geom_point(aes(x = col_group, y = row_group, color = n, size = n), stroke = 0, shape = 15) +
  scale_size(range = c(1,3.4), breaks = c(100,500,1000,1500)) +
  scale_color_gradient(low = "#6baed6", high = "#08306b") +
  scale_x_continuous(breaks = c(1,4,8,12,16),
                     labels = c("March", "April", "May", "June", "July")) +
  scale_y_discrete(limits = rev) +
  guides(color = "none",
        size = guide_legend(override.aes = list(color = "#2171b5"),
                            title.position = "bottom", title.hjust = 0.5)
        ) +
  facet_grid(species ~ .) + 
  theme(
        axis.text.y = element_blank(),
        axis.text.x = element_text(hjust = 0),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(x = "", y = "", color = "seal-hours",
       size = "seal hours of bio-logger data") +
  ggtitle("Density of haul-out behavior observations varies from March through July",
          subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")
```

Haul-out behavior data was recorded and transmitted via the Argos satellite
network as hourly percent-dry timelines. For each hour of a day, the wet/dry
sensor was polled by the tag firmware every few seconds and a percent of the
hour in a dry state was calculated (Figure \@ref(fig:examplePlot)). On board the
bio-logger, hourly percent-dry data were rounded to the nearest 10 percent
inclusive of 0 and 100 percent and additional values at 3 and 98 percent. This
compression resulted in additional data transmission as each message consisted
of two complete 24-hour records. Bio-loggers were, generally, deployed and
programmed in a manner to maximize data transmission during the spring pupping
and molting period, however, hourly percent-dry data for days during a
deployment were not successfully transmitted. This is due to a variety of
factors including satellite coverage, tag availability (i.e. tags mounted to the
rear flipper are not available to the surface while the seal is at sea), tag
performance, duty cycling, and extra-terrestrial atmospheric interference.

@@

p <- p + facet_grid(year~month, scales = "free_x")
p <- p + scale_x_continuous(breaks = c(5,15,25), expand = c(0,0)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  theme(panel.spacing = unit(0,"line")) +
  ggtitle("Observation records are incomplete because not all data are transmitted",
          subtitle = "percent-dry data for a single ribbon seal across two years") +
  xlab("day of month") + ylab("hour (UTC)") 
p <- p + 
  geom_point(aes(x = day, y = hour), data = missing_hours, 
             color = "black", shape = 4, size=0.8)
p
```

Of key interest in this study is the relationship between haul-out behavior and
weather covariates that vary with space and time. We explored the use of a
continuous-time correlated random walk [@devins.johnson2008] movement model to
predict locations at specific times. However, the sparse nature of data from
some bio-loggers, especially those mounted to the rear flipper, resulted in poor
modeling performance or convergence issues. For this study, we calculated a
weighted average daily location where the inverse of the estimated Argos or
FastLoc GPS location error was used for the weight. Each Argos location estimate
was assigned an error radius based on either the categorical location quality (
_3_=250m, _2_=500m, _1_=1500m, _0_=2500m, _A_=2500m, _B_=2500m) or, when
available, the estimated error radius from the Argos Kalman filter algorithm.
Location estimates from FastLoc GPS were all assigned an error radius of 50m.
Any days where haul-out observations were present but location data were missing
we used the last calculated weighted average daily location, and any days where
the location intersected with land were removed from the data set. We recognized
that bearded and spotted seals haul out on land. However, assessing the
relationship between haul-out behavior and weather covariates and their
availability for aerial surveys on land was outside the scope of this study.
Additionally, any daily locations on land are likely more reflective of
coordinate averaging and measurement error instead of actual use of coastal
features. Figure \@ref(fig:dataMap) shows the spatial distribution weighted
locations with available haul-out behavior data for each species across the
study area.

@@

In addition to sex and age class, we analyzed variables that might help explain
variation in haul-out probabilities. These included day-of-year (for seasonal
effects) and local solar hour (for diurnal effects). Solar hour was calculated
using the {solaR} package [@perpinan2012] within the R statistical environment
[@rcoreteam2021] using the daily locations. We included several weather
variables shown to affect haul-out behavior in other Arctic pinnipeds
[@reder2003; @udevitz2009; @perry2017]. In particular, we linked the weighted
average daily locations from our bio-loggers to key weather values from the
North American Regional Reanalysis (NARR) model produced by the National Centers
for Environmental Prediction [@mesinger2006]. The NARR model assimilates
observational data to produce a long-term picture of weather over North America.
Numerous weather variables are made available across the region 8 times daily.
For this study, NARR weather values were subset to the extent of our study area
over the Bering and Chukchi seas at 3-hr intervals based on the native grid
resolution of 32 km (1024 sq. km). The following meteorological variables were
interpolated and assigned to seal locations using a bilinear method: 1) air
temperature at 2m above the earth's surface, 2) wind consisting of northerly and
easterly vector components converted to wind speed using the Euclidean norm, 3)
barometric pressure at sea level, and 4) precipitation (Table
\@ref(tab:tblCovariates)).

@@

covariate_tbl <- tibble::tribble(
  ~Covariate, ~Type, ~Source, ~Description,
  "Age-sex class",   "Categorical", "Field Assessment", 
  "young-of-the-year, subadult, adult male and adult female",
  "Hour", "Continuous; Fourier basis", "Bio-logger", 
  "local solar hour using 6 variables of a Fourier-series basis",
  "Day",   "Continuous", "Bio-logger", 
  "linear, quadratic, and cubic effects of day-of-year",
  "Precip", "Continuous", "NARR", 
  "convective precipitation (kg/m2)",
  "Pressure", "Continuous", "NARR", 
  "atmospheric pressure at sea level (kPa)",
  "Temp", "Continuous", "NARR",
  "air temperatures at 2m above the earth’s surface",
  "Wind", "Continuous", "NARR", 
  "northerly and easterly vector components for wind converted into a single wind speed via the Euclidean norm",
  "Northing", "Continuous", "Bio-logger","latitude divided by the mean latitude across all locations (for bearded seals only)",
  "Year", "Continuous", "Bio-logger", "For the set of models examining inter-annual variation in sea ice use, we fitted models with the addition of year by day-of-year interactions." 
)

covariate_tbl %>% 
  flextable() %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  hline(i=8) %>% 
  set_table_properties(width = .95, layout = "autofit")
```

@@

When exploring the influence of weather, Bearded seal haul-out probability appears most influenced by wind
($F_{`r paste(bearded_type3$wind$Num.df, bearded_type3$wind$Den.df, sep=",")`}$
= `r bearded_type3$wind$F.value`; $p$ = `r bearded_type3$wind$Prob.F`)
and temperature
($F_{`r paste(bearded_type3$temp2$Num.df, bearded_type3$temp2$Den.df, sep=",")`}$
= `r bearded_type3$temp2$F.value`; $p$ =
`r bearded_type3$temp2$Prob.F`). Additionally, wind chill was also a
significant influence
($F_{`r paste(bearded_type3$temp2_wind$Num.df, bearded_type3$temp2_wind$Den.df, sep=",")`}$
= `r bearded_type3$temp2_wind$F.value`; $p$ =
`r bearded_type3$temp2_wind$Prob.F`). Any influence of barometric
pressure or precipitation was less apparent.

@@

Compared to ribbon seals, spotted seals showed a longer spring haul-out season
that is less intensely centered on solar noon (Figure \@ref(fig:spottedHOCal)).
Adults of both sexes spend considerable time in April, May and June hauled out.
In contrast to ribbon seals, adult spotted seal males have a more protracted
haul-out season compared to females, and more time out of the water in June and
July (Figures \@ref(fig:ribbonHOCal) and \@ref(fig:spottedHOCal)). As with
ribbon seals, the young-of-the-year records begin after weaning and the model
predictions demonstrate the ontogeny of in-water activities (e.g. diving,
foraging) in May.

@@

We modeled data from bio-loggers deployed on bearded, ribbon, and spotted seals
to examine factors affecting haul-out behavior on sea ice in the Bering and
Chukchi seas. Our analysis shows all three species of seal haul out
progressively more through the spring and peak near mid-May to early June before
declining again. This pattern aligns well with what has been previously
documented and confirms our haul-out data are likely capturing population-level
behavioral patterns. Seals preferentially haul out on ice shortly after solar
noon, which allows seals to maximize absorption of solar radiation thought to
facilitate the molting process [@feltz1966]. Interestingly, bearded seals appear
to have two peaks in haul-out activity within a day, one shortly after solar
noon, and one centered near solar midnight. A similar bi-modal pattern has been
seen in ringed seals [@vonduyke2020] and suggests that bearded and ringed seals
may be operating under different constraints than ribbon and spotted seals.
Bearded and ringed seals are distributed across higher latitudes and the
extended daylight hours may allow more flexibility in optimizing resting periods
with foraging. Other factors such as polar bear predation (which is rare for
ribbon and spotted seals in the Bering Sea) may also explain differing haul-out
patterns. Bearded seals also showed a more protracted season of haul-out
behavior compared to ribbon and spotted seals. This aligns with findings from
Thometz et al. [@thometz2021] who observed a mean molting period of 119&pm;2
days and a relatively stable resting metabolic rate during that time. While
ribbon seals were not considered in the same study, spotted and ringed seals
underwent molt periods of 33&pm;4 and 28&pm;6 days and exhibited increased
resting metabolic rates.

@@

We also investigated the influence of weather on haul-out
probabilities, including wind speed, temperature, barometric pressure,
precipitation, and wind chill. These have been investigated for walruses (e.g.
Udevitz et al. [-@udevitz2009]) and a few select studies of ice-associated seals
[@hamilton2018, @perry2017]. Ribbon seals seemed to be the most influenced by
weather, with wind, temperature, barometric pressure, and precipitation all
being important components of the model. Spotted seals were most affected by
wind and barometric pressures. And, for bearded seals, the model indicated wind
and temperature had the greatest impact. In general, and as might be expected,
seals were more likely to haul-out when daily temperatures were warmer, winds
speeds were lower, barometric pressure was higher, and precipitation was lower.
These weather conditions are likely most beneficial for energetic considerations
and may also provide an optimal situation for predator detection. These results
highlight the importance for haul-out behavior models (and subsequent
calculations of availability during surveys) to incorporate weather covariates.
